|

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
update text 2.3.3

Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely the null hypothesis is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of the null hypothesis despite when majority of the residuals are well-behaved [@cook_applied_1999]. Furthermore, with a sufficiently large sample size, residual-based tests may reject the null hypothesis due to a slight departure that is of little practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.

@@

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004]. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis. Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup. Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$. If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.

@@

Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the actual data plot at position 7 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified. In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.

@@

Data used in the $m - 1$ null plots need to be simulated. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

@@

Further, a visual test can involve $K$ independent observers. Let $D_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot. Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$. Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].

As pointed out by @vanderplas2021statistical, the binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as

@@

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},
\end{equation}
```

@@

Note that Equation \ref{eq:pvalue-beta-binomial} only works with integer $c_i$. For non-integer $c_i$, linear approximation will be applied to calculate the p-value

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial-approx}
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \notin \mathbb{Z}.
\end{equation}
```

<!-- TODO:Introduce the concept of c interesting plot -->

Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be estimated from the survey data. For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected. For higher values of $\alpha$, the distribution of the probability of each plot being selected is more evenly. @vanderplas2021statistical defined that a plot is $c$-interesting if $c$ or more participants selected the plot as the most different. Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as

@@

@vanderplas2021statistical suggested that $\alpha$ can be estimated using MLE and Equation \ref{eq:c-interesting-expectation}. But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup consists of only null plots, are required.

<!-- But if $\alpha$ is small and only a few null plots in a lineup are attractive, MLE could fail to provide accurate estimates. -->

@@

```{r residual-plot-cubic-heter, fig.cap='Residuals vs. fitted values plot for a classical normal linear regression model. The residuals are produced by fitting a two-predictor multiple linear regression model with data generated from a cubic linear model. From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity. Further, from the outline of the shape, nonlinear patterns exist. Both visual discoveries are evidence against the null hypothesis, though heteroskedasticity actually does not exist in the data generating process. \\label{fig:residual-plot-cubic-heter}', eval = FALSE}

@@

In particular, given 95% confidence level, both tests will reject the null hypothesis in only around 58% of the cases. And in 41% of the cases, the visual test does not reject the null hypothesis but the conventional test does. Since fail to reject the null hypothesis in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of the null hypothesis even though the alternative hypothesis is true. Even if the rejection is accepted, the model violation may not be considered as impactful due to the fact that the departures are not clearly visible.

This makes it sounds like people are arrogantly ignoring the accuracy of conventional tests and trust too much to their visual perception, but it is not. In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a simple but good linear approximation to the data. The rejection of the null hypothesis because of acceptable and negligible residual departure is not practically meaningful and useful. This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.

@@

In terms of the heteroskedasticity model, the power of Breusch-Pagan test is also almost always greater than the power of visual test. For $0 \leq log_e(E) \leq 2$, where the visual feature is nearly unobservable from the example residual plots and the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting the null hypothesis.

@@

The power of all inappropriate tests except for F-test shows improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of F-test remains at around 5% in Figure \ref{fig:heterpower} since there is no non-linear term leave out in the heteroskedasticity model and 
the null hypothesis of the F-test will be always satisfied.

>>>>

Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis $H_0$ for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely $H_0$ is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999]. Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.

@@

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004]. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with $H_0$. Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup. Under $H_0$, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$. If we reject $H_0$ as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.

@@

Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the actual data plot at position 7 is identifiable, then it is evidence for the rejection of $H_0$ that the regression model is correctly specified. In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.

@@

Data used in the $m - 1$ null plots need to be simulated. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggest, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

@@

Further, a visual test can involve $K$ independent observers. Let $X_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot. Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under $H_0$, $X \sim \mathrm{Binom}_{K,1/m}$. Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as $P(X \geq x) = 1 - F(x) + f(x)$, where $F(.)$ is the cumulative distribution function, $f(.)$ is the probability mass function and $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013].

As point out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. @vanderplas2021statistical summarizes three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design. For Scenario 3, @vanderplas2021statistical model the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as

@@

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
\end{equation}
```

@@

Note that Equation \ref{eq:pvalue-beta-binomial} only works with non-negative integer $c_i$. For non-negative real number $c_i$, linear approximation will be applied to calculate the p-value

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial-approx}
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
\end{equation}
```

where $P(C \geq \lceil c_i \rceil)$ is calculated using Equation \ref{eq:pvalue-beta-binomial} and $P(C = \lfloor c_i \rfloor)$ is calculated by

\begin{equation}
P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
\end{equation}

Besides, the parameter $\alpha$ used in Equation \ref{eq:pvalue-beta-binomial} is usually unknown and hence needs to be estimated from the survey data. For low values of $\alpha$, only a few plots are attractive to the observers and tend to be selected. For higher values of $\alpha$, the distribution of the probability of each plot being selected is more even. @vanderplas2021statistical define that a plot is $c$-interesting if $c$ or more participants select the plot as the most different. Given the definition, The expected number of plots selected at least $c$ times, $E[Z_c]$, is calculated as

@@

@vanderplas2021statistical suggest that $\alpha$ can be estimated using MLE with Equation \ref{eq:c-interesting-expectation}. But for precise estimate of $\alpha$, additional responses to Rorschach lineups, which is a type of lineup consists of only null plots, are required.

@@

```{r residual-plot-cubic-heter, fig.cap='Residuals vs. fitted values plot for a classical normal linear regression model. The residuals are produced by fitting a two-regressor multiple linear regression model with data generated from a cubic linear model. From the residual plot, "butterfly shape" can be observed which generally would be interpretd as evidence of heteroskedasticity. Further, from the outline of the shape, nonlinear patterns exist. Both visual discoveries are evidence against $H_0$, though heteroskedasticity actually does not exist in the data generating process. \\label{fig:residual-plot-cubic-heter}', eval = FALSE}

@@

In particular, given 95% confidence level, both tests will reject $H_0$ in only around 58% of the cases. And in 41% of the cases, the visual test does not reject $H_0$ but the conventional test does. Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not fully believe in the rejection of $H_0$ even though the alternative hypothesis is true. Even if the rejection is accepted, the model violation may not be considered as impactful due to the fact that the departures are not clearly visible.

This makes it sounds like people are arrogantly ignoring the accuracy of conventional tests and trust too much to their visual perception, but it is not. In fact, the sensitivity of the conventional test could sometimes distract and discourage analysts from finding a simple but good linear approximation to the data. The rejection of $H_0$ because of acceptable and negligible residual departure is not practically meaningful and useful. This also partly explains why abundance of literature suggesting the use of data plot in regression diagnostics.

@@

In terms of the heteroskedasticity model, the power of Breusch-Pagan test is also almost always greater than the power of visual test. For $0 \leq log_e(E) \leq 2$, where the visual feature is nearly unobservable from the example residual plots and the power curve of the visual test remains at a low level, the Breusch-Pagan test still have a decent amount of chance of rejecting $H_0$.

@@

The power of all inappropriate tests except for F-test shows improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of F-test remains at around 5% in Figure \ref{fig:heterpower} since there is no non-linear term leave out in the heteroskedasticity model and 
$H_0$ of the F-test will be always satisfied.
